Parsing Postfix log files

### 2008-10-31 doc/logparser-introduction.tex
Finished corrections; 112 XXX notes.

Section~\ref{rules} discusses the parsing rules in detail, explaining the
purpose and usage of each field in a rule, referring to an example rule and
sample data it matches successfully against.  The pros and cons of
overlapping rules are considered, including techniques for detecting
unintentional overlaps.  Rule efficiency concerns are discussed, in
particular the optimisations used by the algorithm.  The section concludes
with a description of using the tools provided with the parser to generate
new rules (specifically the \regex{} in each rule) from unparsed log lines.

>>>>

Section~\ref{rules} discusses the parsing rules in detail, explaining the
purpose and usage of each field in a rule, referring to an example rule and
sample data it matches successfully against.  The pros and cons of
overlapping rules are considered, including techniques for detecting
unintentional overlaps.  Rule efficiency concerns are discussed, in
particular the optimisations used by the algorithm.  The section concludes
with a description of using the tools provided with the parser to generate
new rules (specifically the regex in each rule) from unparsed log lines.

### 2008-10-31 doc/logparser-other-parsers-review.tex
Finished corrections; 112 XXX notes.

\LMA{} requires major changes to the parser code to parse new log lines or
extract additional data.  The code is structured as a long series of blocks
that each handle all lines matching a single \regex{}, so parsing new log
lines requires modifying an existing \regex{} or carefully inserting a new
block in the correct place; extracting extra data will require modifying
multiple blocks, \regexes{}, or both.

@@

% Literature review notes:
%
% Hard-coded parsing, requiring code changes to add more.  Attempts to
% correlate log lines, saves data to database for data mining purposes.
% Hard to extend/expand/understand.  Appears to only save: date and hour,
% \DNS{} name and \IP{} address host, mail server \IP{} address, sender,
% receiver and e-mail status (sent, rejected).  Undocumented schema.
% Design decision to use \CSV{} as an intermediate format between the log
% file and the database seems to have been restrictive.  Appears to require
% a queueid but majority of log entries (e.g.\ rejections) lack a queueid.
% Supports whitelisting \IP{} addresses when parsing logs, but whitelisting
% when generating reports/data mining would be preferable.  Supporting
% Berkeley DB is probably limiting the software - an example is the
% difficulty in searching a pipe-delimited string, so they have
% re-implemented foreign keys with tables keyed by ip address etc.\
% pointing at the main table - this also will not scale well.  There does
% not appear to be any attempt to deal with the complications I have
% encountered: their parsing is not detailed enough to encounter them.  It
% does not run properly; does not create any output; throws up errors.

@@

It also counts the number of log lines parsed and skipped.  Log lines are
correlated based on the queueid (referred to as sessionname [sic] within
Logrep), but not by \pid{}.  The parsing is error prone: empty fields are
saved when the log line does not match the \regex{}, though it appears that
they will not overwrite existing data.  Most notably rejections are
completely ignored, making it unsuitable for the purposes of this project.
It does not try to address any of the complications in
\sectionref{complications} except for correlating by queueid.

>>>>

\LMA{} requires major changes to the parser code to parse new log lines or
extract additional data.  The code is structured as a long series of blocks
that each handle all lines matching a single regex, so parsing new log
lines requires modifying an existing regex or carefully inserting a new
block in the correct place; extracting extra data will require modifying
multiple blocks, regexes, or both.

@@

% Literature review notes:
%
% Hard-coded parsing, requiring code changes to add more.  Attempts to
% correlate log lines, saves data to database for data mining purposes.
% Hard to extend/expand/understand.  Appears to only save: date and hour,
% DNS name and \IP{} address host, mail server \IP{} address, sender,
% receiver and e-mail status (sent, rejected).  Undocumented schema.
% Design decision to use \CSV{} as an intermediate format between the log
% file and the database seems to have been restrictive.  Appears to require
% a queueid but majority of log entries (e.g.\ rejections) lack a queueid.
% Supports whitelisting \IP{} addresses when parsing logs, but whitelisting
% when generating reports/data mining would be preferable.  Supporting
% Berkeley DB is probably limiting the software - an example is the
% difficulty in searching a pipe-delimited string, so they have
% re-implemented foreign keys with tables keyed by ip address etc.\
% pointing at the main table - this also will not scale well.  There does
% not appear to be any attempt to deal with the complications I have
% encountered: their parsing is not detailed enough to encounter them.  It
% does not run properly; does not create any output; throws up errors.

@@

It also counts the number of log lines parsed and skipped.  Log lines are
correlated based on the queueid (referred to as sessionname [sic] within
Logrep), but not by \pid{}.  The parsing is error prone: empty fields are
saved when the log line does not match the regex, though it appears that
they will not overwrite existing data.  Most notably rejections are
completely ignored, making it unsuitable for the purposes of this project.
It does not try to address any of the complications in
\sectionref{complications} except for correlating by queueid.

### 2008-10-31 doc/logparser-results.tex
Finished corrections; 112 XXX notes.

Before the perl interpreter attempts to match a \regex{} against a piece of
text, the \regex{} is compiled into an internal representation and
optimised to improve the speed of matching.  This compilation and
optimisation takes CPU time ---in many cases it takes far more CPU time
than the actual matching.  If the interpreter is certain that a \regex{}
will not change it will automatically cache the compilation results for
later use.   The results of compiling a dynamically generated \regex{} can
be cached and used in preference to the original \regex{}, but it is the
responsibility of the programmer to do this.  The Postfix parser loads the
rules it uses from a database, and thus each rule's \regex{} is compiled
and the result cached to reduce parsing time.

\Graphref{Increase in parsing time when not caching compiled regexes graph}
shows the impact that not caching compiled \regexes{} has on parser
performance: on typical log files the parsing time when not caching
compiled \regexes{} is 400--600\% of the parsing time when caching.
Caching compiled \regexes{} is probably the single most effective
optimisation possible in the parser's implementation, and was quite simple
to implement.

@@

Two large dips can be seen in \graphref{Increase in parsing time when not
caching compiled regexes graph} at log files 22 and 62--68, corresponding
to the spikes in log file size in \graphref{parsing time vs file size vs
number of lines graph}.  The distribution of log lines across rules when
there is a mail loop is much different to the norm, and the average number
of rules consulted per log line is much lower: this results in far fewer
\regex{} compilations per line when there is a mail loop, and a
corresponding decrease in the average parsing time for a single log line.
The outcome is that caching compiled \regexes{} is proportionally less
important when the log file contents were created by a mail loop.  The
increases in parsing time when not caching compiled \regexes{} for
different groups of log files are summarised in \tableref{Increase in
parsing time when not caching compiled regexes table}.

@@

This section began with parser scalability, showing the linear relationship
between parsing time and input size.  Demonstrates the effect of rule
ordering on parsing time, and the unexpected consequences of specific
inputs.  The necessity of caching compiled \regexes{} is attested to by the
third group of graphs, where the difference between caching and not caching
is staggering.  The penultimate section contains a breakdown of the rule
hits accumulated during a single test run of the parser.  Miscellaneous
graphs expected to be useful are collected in the final section.

@@

Full coverage of log lines is easy to achieve yet hard to maintain.  It is
easy to achieve full coverage for a limited set of log files (at the time
of writing the parser is tested with \numberOFrules{} rules, fully parsing
\numberOFlogFILESall{} contiguous log files from Postfix 2.2 through to
Postfix 2.5), and new rules are easy to add.  Maintaining full coverage is
hard because other servers have different restrictions with custom
messages, \DNSBL{} messages change over time, major releases of Postfix
change warning messages (usually adding more information), etc.,\ so over
time the log lines drift and change.  \Graphref{rule hits graph} shows the
number of hits for each rule over all \numberOFlogFILES{} log files.  The
number of hits per rule is quite unevenly spread, resembling a Power Law
distribution~\cite{powerlaw} --- it is obvious that a small number of rules
match the vast majority of the lines, and more than half the rules match
fewer than 100 times.

@@

Warnings are issued for any log lines that are not parsed; no warnings are
issued for unparsed log lines while testing with the \numberOFlogFILES{}
test log files, so it can be safely concluded that there are zero false
negatives.  False positives are harder to quantify: short of examining each
of the 60,721,709 log lines and determining which \regex{} parsed it, there
is no way to be sure that every line was parsed by the correct \regex{},
making it impossible to quantify the false positive rate; however a random
sample of 6,039 log lines was parsed and the results checked manually to
ensure that the correct \regex{} parsed each log line.\footnote{Each log
line was examined and the correct \regex{} identified from the
\numberOFrules{} rules in the database; the correct \regex{} was then
compared to the \regex{} that was used by the parser.}  The sample was
generated by running the following command:

>>>>

Before the perl interpreter attempts to match a regex against a piece of
text, the regex is compiled into an internal representation and
optimised to improve the speed of matching.  This compilation and
optimisation takes CPU time ---in many cases it takes far more CPU time
than the actual matching.  If the interpreter is certain that a regex
will not change it will automatically cache the compilation results for
later use.   The results of compiling a dynamically generated regex can
be cached and used in preference to the original regex, but it is the
responsibility of the programmer to do this.  The Postfix parser loads the
rules it uses from a database, and thus each rule's regex is compiled
and the result cached to reduce parsing time.

\Graphref{Increase in parsing time when not caching compiled regexes graph}
shows the impact that not caching compiled regexes has on parser
performance: on typical log files the parsing time when not caching
compiled regexes is 400--600\% of the parsing time when caching.
Caching compiled regexes is probably the single most effective
optimisation possible in the parser's implementation, and was quite simple
to implement.

@@

Two large dips can be seen in \graphref{Increase in parsing time when not
caching compiled regexes graph} at log files 22 and 62--68, corresponding
to the spikes in log file size in \graphref{parsing time vs file size vs
number of lines graph}.  The distribution of log lines across rules when
there is a mail loop is much different to the norm, and the average number
of rules consulted per log line is much lower: this results in far fewer
regex compilations per line when there is a mail loop, and a
corresponding decrease in the average parsing time for a single log line.
The outcome is that caching compiled regexes is proportionally less
important when the log file contents were created by a mail loop.  The
increases in parsing time when not caching compiled regexes for
different groups of log files are summarised in \tableref{Increase in
parsing time when not caching compiled regexes table}.

@@

This section began with parser scalability, showing the linear relationship
between parsing time and input size.  Demonstrates the effect of rule
ordering on parsing time, and the unexpected consequences of specific
inputs.  The necessity of caching compiled regexes is attested to by the
third group of graphs, where the difference between caching and not caching
is staggering.  The penultimate section contains a breakdown of the rule
hits accumulated during a single test run of the parser.  Miscellaneous
graphs expected to be useful are collected in the final section.

@@

Full coverage of log lines is easy to achieve yet hard to maintain.  It is
easy to achieve full coverage for a limited set of log files (at the time
of writing the parser is tested with \numberOFrules{} rules, fully parsing
\numberOFlogFILESall{} contiguous log files from Postfix 2.2 through to
Postfix 2.5), and new rules are easy to add.  Maintaining full coverage is
hard because other servers have different restrictions with custom
messages, \DNSBL{} messages change over time, major releases of Postfix
change warning messages (usually adding more information), etc.,\ so over
time the log lines drift and change.  \Graphref{rule hits graph} shows the
number of hits for each rule over all \numberOFlogFILES{} log files.  The
number of hits per rule is quite unevenly spread, resembling a Power Law
distribution --- it is obvious that a small number of rules match the vast
majority of the lines, and more than half the rules match fewer than 100
times.

@@

Warnings are issued for any log lines that are not parsed; no warnings are
issued for unparsed log lines while testing with the \numberOFlogFILES{}
test log files, so it can be safely concluded that there are zero false
negatives.  False positives are harder to quantify: short of examining each
of the 60,721,709 log lines and determining which regex parsed it, there
is no way to be sure that every line was parsed by the correct regex,
making it impossible to quantify the false positive rate; however a random
sample of 6,039 log lines was parsed and the results checked manually to
ensure that the correct regex parsed each log line.\footnote{Each log
line was examined and the correct regex identified from the
\numberOFrules{} rules in the database; the correct regex was then
compared to the regex that was used by the parser.}  The sample was
generated by running the following command:

### 2008-10-31 doc/logparser.tex
First half of the corrections to section 5.
Fixed underfull hbox warnings. Dealt with more corrections; nearly finished (if I ignore the 103 XXX notes).
Finished corrections; 112 XXX notes.

\renewcommand{\refname}{Bibliography}
% New formatting commands.

@@

@@

%\bibliographystyle{logparser-bibliography-style}
\bibliographystyle{sgai-2008-elsevier}
\bibliography{logparser-bibliography}
\label{bibliography}

% Redefine the command used to produce the glossary title, because the
% default command produces an unnumbered section whereas I want a numbered
% section.
\renewcommand{\glossarytitle}{\section{Glossary}\label{Glossary}}
\printglossary{}
% Redefine the command a second time to produce acronyms instead of a
% glossary.
\renewcommand{\glossarytitle}{\section{Acronyms}\label{Acronyms}}
\printacronym{}

>>>>

% Show the citation key (e.g. slct-paper) in the margin beside the
% reference.
\usepackage[nomsgs]{refcheck}
% Show where references are used.
% Neither work.
%\usepackage{citeref}
%\usepackage{backref}
\renewcommand{\refname}{Bibliography}
% Add space before urls to quell underfull hbox warnings.
\newcommand{\urlprefix}[0]{\hfill{}}
% New formatting commands.

@@

XXX WOULD TABLES LOOK BETTER IF CAPTIONS WERE LEFT-ALIGNED\@?

XXX SWITCH TO NEW GLOSSARY PACKAGE\@.

XXX IS IT POSSIBLE TO HAVE ACRONYMS THAT ARE NOT INCLUDED IN THE ACRONYM
LIST\@?

XXX REDUCE THE LIST OF ACRONYMS\@.

XXX CHECK THAT I SAY log line AND log file EVERYWHERE\@.

XXX CHECK YEARS FOR BIBLIOGRAPHY ENTRIES\@.

XXX IMPROVE BIBLIOGRAPHY APPEARANCE\@: NAMES LOOK UGLY (E.G.\ P.\ \empty{}.\
Porters), URL SPACING IS MESSED UP --- URLS \\SHOULD PROBABLY BE LAST, ON THEIR
OWN LINE\@.

@@

%\bibliographystyle{logparser-bibliography-style}
\bibliographystyle{sgai-2008-elsevier}
\bibliography{logparser-bibliography}

% Redefine the command used to produce the glossary title, because the
% default command produces an unnumbered section whereas I want a numbered
% section.
XXX MOVE POSTFIX DAEMONS INTO A SEPARATE SECTION\@; NEED TO SWITCH TO NEW
GLOSSARY PACKAGE FIRST\@.
\renewcommand{\glossarytitle}{\section{Glossary}\label{Glossary}}
\printglossary{}
% Redefine the command a second time to produce acronyms instead of a
% glossary.
\renewcommand{\glossarytitle}{\section{Acronyms}\label{Acronyms}}
\printacronym{}
